\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-05-05 doc/paper.tex
document `PotentialLSPParticles' variable

\section{Program Summary}
\noindent{\em Program title:} \fs\\ {\em Program obtainable from:}
         {\tt http://flexiblesusy.hepforge.org/}\\ {\em Distribution
           format:}\/ tar.gz\\ {\em Programming language:} {\tt
           C++}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
           system:}\/ Tested on Linux 3.x\\ {\em Word size:}\/ 64
         bits\\ {\em External routines:}\/ SARAH 4.0.4, Boost library,
         lapack\\ {\em
           Typical running time:}\/ 0.1-0.3 seconds per parameter
         point.\\ {\em Nature of problem:}\/ Determining the mass
         spectrum and mixings for any supersymmetric model. The
         generated code must find simultaneous solutions to
         constraints which are specified at two or more different
         renormalisation scales, which are connected by
         renormalisation group equations forming a large set of
         coupled differential equations. \\ {\em Solution method:}\/
         Nested iterative algorithm and numerical minimisation of the
         Higgs potential.  \\ {\em Restrictions:} The couplings must
         remain perturbative at all scales between the highest and
         lowest boundary condition.  \fs~ assumes that all couplings
         of the model are real (i.e.\ $CP-$conserving)?  Due to the
         modular nature of the generated code adaption and extension
         to overcome restrictions in scope is quite straight forward.

@@

\section{The Program}
\label{Sec:Program}
\fs is a MATHEMATICA package designed to create a fast and easily
adaptable spectrum generator in C++ for any SUSY model and some
non-supersymmetric models {\color{red} can we?}.  \fs makes use of the
existing \sarah package \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} to obtain one-loop self energies, one-loop
tadpoles corrections and two-loop renormalisation group
equations(RGEs) in any model.  \fs then converts these algerbraic
expressions from \sarah into C++ code using the C++11 standard. \fs
uses some parts of SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen
library \cite{eigen}, as well as the gnu scientific library and the
boost library to create numerical routines which then use these
routines to solve the RGEs and EWSB conditions which will be used in
the spectrum generator. The main solver routine of \fs then calls
these routines to find simultaneous solutions for the EWSB, low energy
data, and other, user supplied, boundary conditions (e.g. the usual
mSUGRA universality conditions) and calculates the mass spectrum.

 The modular structure of the generated code allows for any individual
 components to be replaced with an alternative if available.  Indeed
 \fs has been intentionally designed to grow as alternative solvers
 and calculators are added and a subsequent release with an
 alternative lattice solver is already planned.{\color{red} Or Indeed
   flexible SUSY is distributed with two different main solvers, the
   two-scale solver based on the approach of \cite{} and designed for
   speed and the a lattice solver designed to solve tricky cases where
   the fixed point iteration of the two-scale approach fails to
   converge.}  In addition if one already has an existing code for
 some other purpose and simply wishes to improve upon this by adding,
 e.g, new RGEs or self energies the modular fraework makes this fairly
 straightward{\color{red} Is this true?  because it should be...}

@@

OnlyLowEnergyFlexibleSUSY = False; (* default *)
\end{lstlisting}
%
The variable \code{FSModelName} contains the name of the \fs model.

@@

>>>>

\section{Program Summary}
\noindent{\em Program title:} \fs\\ {\em Program obtainable from:}
         {\tt http://flexiblesusy.hepforge.org/}\\ {\em Distribution
           format:}\/ tar.gz\\ {\em Programming language:} {\tt
           C++}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
           system:}\/ Tested on Linux 3.x\\ {\em Word size:}\/ 64
         bits\\ {\em External routines:}\/ SARAH 4.0.4, Boost library,
         lapack\\ {\em
           Typical running time:}\/ 0.1-0.3 seconds per parameter
         point.\\ {\em Nature of problem:}\/ Determining the mass
         spectrum and mixings for any supersymmetric model. The
         generated code must find simultaneous solutions to
         constraints which are specified at two or more different
         renormalisation scales, which are connected by
         renormalisation group equations forming a large set of
         coupled differential equations. \\ {\em Solution method:}\/
         Nested iterative algorithm and numerical minimisation of the
         Higgs potential.  \\ {\em Restrictions:} The couplings must
         remain perturbative at all scales between the highest and
         lowest boundary condition.  \fs~ assumes that all couplings
         of the model are real (i.e.\ $CP-$conserving). Due to the
         modular nature of the generated code adaption and extension
         to overcome restrictions in scope is quite straight forward.

@@

\section{Overview of the Program}
\label{Sec:Program}
\fs is a MATHEMATICA package designed to create a fast and easily
adaptable spectrum generator in C++ for any SUSY model and some
non-supersymmetric models.  The user specifies the model by giving the superfield content, superpotential, gauge symmetries and mass mixings in the  
\sarah model files, examples of which can be found in the \code{Models}
directory of \sarah and the \code{sarah} directory of FlexibleSUSY.  The user
then also specifies the boundary conditions on the model parameters in a separate \code{FlexibleSUSY.m} file, examples of which can be found in the \code{model_files} directory. A detailed guide on how to write the FlexibleSUSY.m model is given in section \ref{Sec:modfile}.

  \fs makes use of the existing \sarah package
  \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
  to obtain tree level expressions for the mass matrices and
  electroweak symmetry breaking conditions, one-loop self energies,
  one-loop tadpoles corrections and two-loop renormalisation group
  equations(RGEs) for the model specified in the model
  files. Additional corrections which have been calculated elsewhere,
  such as two-loop corrections to the Higgs masses\footnote{By default FlexibleSUSY has two loop corrections to the Higgs masses for the MSSM\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}  and NMSSM \cite{Degrassi:2009yq} in FORTRAN files supplied by Pietro Slavich. These are the same corrections which are implemented in many of the public spectrum generators.} may be added in
  file specified by the user.

 \fs then
converts these algerbraic expressions from \sarah into C++ code using
the C++11 standard. \fs uses some parts of
SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen library
\cite{eigen}, as well as the gnu scientific library and the boost
library to create numerical routines which solve the RGEs and EWSB
conditions.  The main solver routine of \fs then calls these routines
to find simultaneous solutions for the EWSB, low energy data, and
other, user supplied, boundary conditions (e.g. the usual mSUGRA
universality conditions) and calculates the mass spectrum.

 The modular structure of the generated code allows for any individual
 components to be replaced with an alternative if available.  Indeed
 \fs has been intentionally designed to grow as alternative solvers
 and calculators are added and a subsequent release with an
 alternative lattice solver is already planned.  In addition if one
 already has an existing code for some other purpose and simply wishes
 to improve upon this by adding, e.g, new RGEs or self energies the
 modular framework makes this straightforward by adding or
 interfacing with the appropriate routine.

@@

OnlyLowEnergyFlexibleSUSY = False; (* default *)

PotentialLSPParticles = { Chi, Cha, Glu, Sv, Su, Sd, Se };
\end{lstlisting}
%
The variable \code{FSModelName} contains the name of the \fs model.

@@

\fs can create the helper function \code{get_lsp()}, which finds the
lightest supersymmetric particle (LSP).  To let this function be
created by \fs, the model file variable \code{PotentialLSPParticles}
must be set to a list of SUSY particles which are potential LSPs.  In
the model file example above, the particles \code{Chi}, \code{Cha},
\code{Glu}, \code{Sv}, \code{Su}, \code{Sd}, \code{Se} (neutralino,
chargino, gluino, sneutrino, up-type squark, down-type squark,
selectron) are considered to be LSPs.
